### 2018-10-30 anzjs2018/anzjs2018.tex
edits

\documentclass[times, doublespace]{anzsauth}
\usepackage{moreverb}
\usepackage{url}
\usepackage{grffile}
\usepackage{lineno}
% \usepackage{lmodern}
\usepackage{lipsum}
\usepackage[UKenglish]{isodate}
\usepackage{float}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
% \newfloat{lstfloat}{htbp}{lop}
% \floatname{lstfloat}{Listing}
% \def\lstfloatautorefname{Listing}

% \doublespace

\input{../glossary.tex}

\newcommand{\bv}{\boldsymbol{v}}
\newcommand{\bz}{\boldsymbol{z}}
\newcommand{\bw}{\boldsymbol{w}}
\newcommand{\bR}{\boldsymbol{R}}
\newcommand{\bT}{\boldsymbol{T}}
\newcommand{\br}{\boldsymbol{r}}
\newcommand{\btheta}{\boldsymbol{\theta}}

\newcommand{\rt}{real-time\ }

\def\volumeyear{2018}
\linenumbers

\begin{document}
\cleanlookdateon
\runningheads{Modeling transit vehicles in real-time}{TOM~ELLIOTT AND THOMAS~LUMLEY}
\title{A real-time particle filter for modeling travel times of transit vehicles through a GTFS-based road network}
\author{Tom Elliott\corrauth~and~Thomas~Lumley}
\affiliation{University of Auckland}
\address{
    Department of Statistics, University of Auckland, 
    Auckland 1142, New Zealand\\
    Email: \texttt{tom.elliott@auckland.ac.nz}
}

\begin{abstract}
Predicting the arrival time of a transit vehicle involves not only
knowledge of its current position and schedule adherance,
but also information about traffic conditions along the remainder of its route.
Road networks are dynamic, 
and can change from free-flowing to highly congested very quickly,
which can effect the arrival time of transit vehicles, 
buses in particular,
% This in turn leads to commuter frustration when they have to wait longer than
% initially anticipated,
so any reliable prediction framework needs to account for \rt traffic conditions.
In this paper,
we propose a particle filter that can be used to estimate travel times
along roads within a transit network based solely on GTFS data.
Due to the \rt nature of the problem and computational demands of such an approach,
we demonstrate the faesibility of our program, 
written in the R programming language and making extensive use of C++.
The application provides the basis from which future work will be based,
most notably for combining the \rt travel times with historical data and a 
formal road network model to make reliable arrival time predictions.
% Of course, the major constraint on any prediction framework is going to be
% computational efficiency.
% Due to the realtime nature of the problem, predictions need to be 
% available as soon as possible after observing the vehicles' positions,
% which are updated approximately every 30~seconds.
% In this paper, we describe our C++ implementation of a particle filter to estimate
% vehicle state in combination with a network state model,
% allowing us to model transit vehicle flow through the network.
Development and testing has taken place here in Auckland, New Zealand,
although we expect our methods are general enough to be applied to any
transit network using GTFS.
\end{abstract}

\keywords{particle filter; transit modeling;
          transit networks; applications; gtfs}

\maketitle
\input{sections/01_introduction}
\input{sections/02_gtfs}
\input{sections/03_models}
\input{sections/04_implementation}
% \input{sections/05_results}
\input{sections/06_discuss}
\bibliographystyle{anzsj}
\bibliography{../reflist.bib}

\end{document}