### 2018-10-31 anzjs2018/anzjs2018.tex
More edits

\documentclass[times, doublespace]{anzsauth}
\usepackage{moreverb}
\usepackage{url}
\usepackage{grffile}
\usepackage{lineno}
% \usepackage{lmodern}
\usepackage{lipsum}
\usepackage[UKenglish]{isodate}
\usepackage{float}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
% \newfloat{lstfloat}{htbp}{lop}
% \floatname{lstfloat}{Listing}
% \def\lstfloatautorefname{Listing}

\doublespace

\input{../glossary.tex}

\newcommand{\bv}{\boldsymbol{v}}
\newcommand{\bz}{\boldsymbol{z}}
\newcommand{\bw}{\boldsymbol{w}}
\newcommand{\bR}{\boldsymbol{R}}
\newcommand{\bT}{\boldsymbol{T}}
\newcommand{\br}{\boldsymbol{r}}
\newcommand{\btheta}{\boldsymbol{\theta}}

\newcommand{\rt}{real-time\ }

\def\volumeyear{2018}
\linenumbers

\begin{document}
\cleanlookdateon
\runningheads{Modeling transit vehicles in real-time}{TOM~ELLIOTT AND THOMAS~LUMLEY}
\title{A real-time particle filter for modeling travel times of transit vehicles through a GTFS-based road network}
\author{Tom Elliott\corrauth~and~Thomas~Lumley}
\affiliation{University of Auckland}
\address{
    Department of Statistics, University of Auckland, 
    Auckland 1142, New Zealand\\
    Email: \texttt{tom.elliott@auckland.ac.nz}
}
\ack{
    We would like to thank Auckland Transport for providing free,
    open access to the \rt data; and the Center for e-Research at the
    University of Auckland for providing and maintaining the virtual
    machine used in our work.
}

\begin{abstract}
Predicting the arrival time of a transit vehicle involves not only
knowledge of its current position and schedule adherance,
but also of traffic conditions along the remainder of the route.
Road networks are dynamic, 
and can change from free-flowing to highly congested very quickly,
which impacts the arrival time of transit vehicles,
particularly buses which often share the road with other vehicles,
so reliable predictions need to account for \rt and future traffic conditions.
The first step in this process is to construct a framework with which road state
(traffic conditions) can be estimated using \rt transit vehicle position data.
We propose a framework which implements a vehicle model using a particle filter
to estimate travel times along roads,
used by a second model to estimate \rt traffic conditions which can
later be used for arrival time prediction.
We demonstrate the \rt faesibility of such an approach in \rt,
and detail our implementation in R and C++.
Development and testing took place in Auckland, New Zealand,
with future work planned to incorporate historical data into the network model
to improve the reliability of ETAs.
\end{abstract}

\keywords{particle filter; transit modeling;
          transit networks; applications; gtfs}

\maketitle
\input{sections/01_introduction}
\input{sections/02_gtfs}
\input{sections/03_models}
\input{sections/04_implementation}
% \input{sections/05_results}
\input{sections/06_discuss}
\bibliographystyle{anzsj}
\bibliography{../reflist.bib}

\end{document}