### 2019-08-03 anzjs2018/sections/04_implementation.tex
Fix graphs

\section{Real-time implementation}
\label{sec:rt}

The application consists of two components:
the first handles importing static GTFS data and constructing the network,
while the second implements the \rt models and prediction.
We used Rcpp \citep{Rcpp}
to develop the program,
which provides access to R \citep{rcore},
which contains many useful tools for data manipulation and package development,
as well as the speed and memory management capabilities of C++.
The program is implemented in the R package
\verb+transitr+, available on Github (\url{https://github.com/tmelliott/transitr}).
In this section, we discuss the features of the \rt component
and assess its performance 
both with respect to timing and parameter estimation.

The general structure of the \rt component is shown below,
with the bold steps being those discussed in this paper.
\begin{enumerate}
\item Load GTFS data from database
\item Each time new data are received \ldots
\begin{enumerate}
    \item Update or create new vehicle objects from the new data
    \item \textbf{Run particle filter on each vehicle to update or initialise its state}
    \item \textbf{Update state of any roads for which vehicles 
        have completed travel}
    \item Generate ETAs for vehicles using combination of vehicle and network state
    \item Write ETAs to a file for distribution
\end{enumerate}
\end{enumerate}

During the development of the application,
we were primarily concerned with ensuring each component of the program
was as efficient as possible,
allowing ETAs to be generated and distributed fast enough to be feasible in real-time,
with a target of 30~seconds or faster at peak time.
Using C++ provided the memory management control necessary to make this possible.
We also made use of OpenMP for parallelisation,
since each vehicle is modelled independently so it is trivial to scale up without
issues with thread safety.

The number of particles needed per vehicle 
depends on many factors,
so it was necessary to explore the performance of the application
with varying number of particles.
Application performance is also assessed
for a range of fixed model parameter values,
such as system noise and GPS error.
To enable comparisons, we implemented a simulated \rt environment
in which the same subset of real vehicle data from 8~October, 2018
could be analysed using a range of settings.
These simulations were carried out of a Virtual Machine 
with 8~cores and 32~Gb of memory, 
running Ubuntu~16.04 and using R~3.4.1.

\subsection{Program Timings}
\label{sec:timings}

In each iteration, 
the timing of the various program components was recorded.
Since the number of vehicles travelling at any given time changes throughout the day,
we used the average timings over an off-peak 15~minute window.
Figure~\ref{fig:timings} shows the average timings for 
varying numbers of particles, $N$.

The most time consuming component was ETA writing,
which involved summarising the individual ETAs estimated for each particle.
While not discussed here this involved computing quantiles which uses a sorting algorithm
with complexity $O(N \log N)$,
which explains the slightly quadratic trend in the timings.

The next most intensive step was vehicle updating,
which took about 5~seconds for 8000 particles.
The ETA prediction step only took a few seconds,
although this may increase once a more comprehensive model is developed.
On our 8-core virtual machine the entire process is completed within 20~seconds
when 8000~particles are used.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/04_model_results_timing.pdf}
    \caption{
        Application timings were recorded for each component of each
        iteration, and averages over a 15~minute window are shown here.
        The timings shown are \emph{wall time}, the actual time passed,
        not CPU time.
        Steps that only involve particle calculations are approximately
        of complexity $O(N)$,
        however some steps are a little more complex (writing ETAs involves computing quantiles), 
        so there is a slightly exponential trend.
    }
        % The timings for various parts of the application, and overall. %
        % The trend is slightly non-linear due to the complexity of the ETA generation step, %
        % which requires sorting of the particle ETAs to obtain quantile estimates.}
    \label{fig:timings}
\end{figure}

\subsection{Model performance}
\label{sec:model_perf}

To evaluate the performance of the model,
the simulation was run using a range of values of GPS error, $\epsilon$,
and system noise, $\sigma^2$.
Figure \ref{fig:dist_to_route} shows the distribution of the distance
between each observation and the route using a nearest point algorithm,
which led us to use the values $\epsilon \in \{1,2,3,5\}$ for GPS error in meters.
For the system noise we used values of $\sigma^2\in \{1e^{-4},1e^{-3},1e^{-2},0.05\}$,
which correspond to an average vehicle speed variance of between 0.1 and 45~meters per second
over a period of 30~seconds.
For each of combination of these parameter values, including varying $N$,
the following values were calculated:
effective sample size, $N_\text{eff}$;
degeneration rate, the percentage of iterations in which the vehicle was lost
(no plausible particles);
and the (relative) variance of segment travel time estimates.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/04_model_results_dist.pdf}
    \caption{
        The distance between the observation and the nearest point on the shape
        is calculated by finding the shortest distance between point and path.
        This gives an approximate idea of GPS error, however it should be noted
        that many observations are made automatically at way points,
        which accounts for the peak at 1m.
        Note also that these observations have been truncated at 20m.
    }
    \label{fig:dist_to_route}
\end{figure}

Figure~\ref{fig:perf_stats} shows that $N_\text{eff}$ increased
with GPS error and decreased with system noise,
while remaining unaffected by changing $N$.
The degeneration rate decreased with GPS error and increased with $N$,
and was unaffected by system noise.
Large GPS error affects the likelihood,
giving more weight to particles farther from the observation
than does a smaller GPS error.
Conversely, increasing system noise spreads out the particle cloud,
so fewer particles will be near the vehicle.
Thus, these results are not surprising, 
but show that the model is working as expected.

\begin{figure}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/04_model_results_neff.pdf}
    \caption{
        The effective sample size is an approximate measure of the variance
        of the particle weights.
        When all particles have equal weights, $N_\text{eff} = N$,
        while if all the weight is on one particle, $N_\text{eff} = 1$. 
        The figure is subset by system noise, which shows how $N_\text{eff}$ 
        decreases as system noise increases,
        while increasing GPS error (on the $x$-axis) increases $N_\text{eff}$.
        Increasing $N$ has a negative effect on $N_\text{eff}$ but this is
        only evident when $N_\text{eff}$ is larger.
        The error bars show $\pm 2$SE.
    }
    \label{fig:perf_stats}
\end{figure}

% \begin{figure}[tb]
%     \centering
%     \includegraphics[width=0.9\textwidth]{figures/04_model_results_degen.pdf}
%     \caption{
%         The degeneration rate is the percentage of iterations where there were no
%         plausible particles, so the vehicle was ``lost'' by the model.
%         This decreases as GPS error and number of particles increase,
%         but there is no visible effect of system noise.
%         Note, however, that some of these can be attributed to invalid data,
%         for example if the driver has entered the wrong trip direction.
%         The error bars show $\pm 2$SE.
%     }
%     \label{fig:degen_rate}
% \end{figure}

The last measure of performance is the relative variance of segment travel times.
We used relative variance as each segment has a very different travel time distribution,
so to compare the variability between simulations we needed to 
compute the average variance of travel times over all simulations,
taking the average (over road segments) of the ratio between variance for
each individual simulation and variance over all simulations.
If $\tilde z_\ell^{\sigma,\epsilon}$ is a vector of all travel times
along segment $\ell$ during the simulation with GPS error $\sigma$ and system noise $\epsilon$,
then the relative variance $\bar v^{\sigma,\epsilon}$ is calculated by
\begin{equation*}
\begin{split}
v_\ell^{\sigma,\epsilon} &=
\frac{\mathrm{Var}(\tilde z_\ell^{\sigma,\epsilon})}{\mathrm{Var}(\tilde z_\ell)} \\
\bar v^{\sigma,\epsilon} &=
    \frac{1}{L}\sum_{\ell=1}^L v_\ell^{\sigma,\epsilon}
\end{split}
\end{equation*}
Figure~\ref{fig:travel_times} shows that simulations with larger GPS error
resulted in a higher relative variance,
while simulations with more particles had lower relative variance.
There was no visible effect of system noise on the variability of travel times.

The results of the simulations
demonstrate a tradeoff between particle filter performance
($N_\text{eff}$ and degeneration rate) and parameter estimation.
However, until the arrival time prediction model has been completed,
it is difficult to make decisions about the optimal values to use:
we are unable to say whether a higher variation of travel time estimates
is going to have a significant effect on the final ETAs,
particularly when compared to the uncertainty of forecasting road state,
or whether reducing the degeneration rate, 
or increasing $N_\text{eff}$, would be more advantageous.

\begin{figure}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/04_model_results_times.pdf}
    \caption{
        The variability of travel time within a segment increases with GPS error,
        and decreases slightly with increasing number of particles.
        There is no visible effect of system noise on travel time variability.
        The error bars show $\pm 2$SE.
    }
    \label{fig:travel_times}
\end{figure}