### 2019-08-10 anzjs2018/anzjs.tex
Grammar and sentence/paragraph restructuring

\documentclass[times, doublespace]{anzsauth}
\usepackage{moreverb}
\usepackage{url}
\usepackage{grffile}
\usepackage{lineno}
% \usepackage{lmodern}
\usepackage{lipsum}
\usepackage[UKenglish]{isodate}
\usepackage{float}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
% \newfloat{lstfloat}{htbp}{lop}
% \floatname{lstfloat}{Listing}
% \def\lstfloatautorefname{Listing}

% \doublespace
\usepackage{afterpage}

\newcommand{\bv}{\boldsymbol{v}}
\newcommand{\bz}{\boldsymbol{z}}
\newcommand{\bw}{\boldsymbol{w}}
\newcommand{\bR}{\boldsymbol{R}}
\newcommand{\bT}{\boldsymbol{T}}
\newcommand{\br}{\boldsymbol{r}}
\newcommand{\btheta}{\boldsymbol{\theta}}

\newcommand{\rt}{real-time\ }

\def\volumeyear{2018}
% \linenumbers

\begin{document}
\cleanlookdateon
\runningheads{Modelling the travel time of transit vehicles}{TOM~ELLIOTT AND THOMAS~LUMLEY}
\title{
    Modelling the travel time of transit vehicles in \rt through a GTFS-based road network
    using GPS vehicle locations
}

% Modelling transit vehicles to estimate the \rt traffic conditions in a public transport network}
\author{Tom Elliott\corrauth~and~Thomas~Lumley}
\affiliation{University of Auckland}
\address{
    Department of Statistics, University of Auckland, 
    Auckland 1142, New Zealand\\
    Email: \texttt{tom.elliott@auckland.ac.nz}
}
\ack{
    We would like to thank Auckland Transport for providing free,
    open access to the \rt data; and the Center for e-Research at the
    University of Auckland for providing and maintaining the virtual
    machine used in our work.
}

\begin{abstract}
Predicting the arrival time of a transit vehicle involves not only
knowledge of its current position and schedule adherence,
but also traffic conditions along the remainder of the route.
Road networks are dynamic 
and can quickly change from free-flowing to highly congested,
which impacts the arrival time of transit vehicles,
particularly buses which often share the road with other vehicles,
so reliable predictions need to account for \rt and future traffic conditions.
The first step in this process is to construct a framework with which road state
(traffic conditions) can be estimated using \rt transit vehicle position data.
Our proposed framework implements a vehicle model using a particle filter
to estimate road travel times,
which are used in a second model to estimate \rt traffic conditions.
Although development and testing took place in Auckland, New Zealand,
we generalised each component to make the framework compatible 
with other public transport systems around the world.
We demonstrate the \rt feasibility and performance of our approach in real-time,
where a combination of R and C++ was used to obtain the necessary performance results.
Future work will use these estimated traffic conditions 
in combination with historical data to obtain reliable
arrival time predictions of transit vehicles.
\end{abstract}

\keywords{particle filter; transit modeling;
          transit networks; applications; gtfs}

\maketitle
\input{sections/01_introduction}
\input{sections/02_gtfs}
\input{sections/03_models}
\input{sections/04_implementation}
\input{sections/05_discuss}
\bibliographystyle{anzsj}
\bibliography{reflist.bib}

\end{document}