### 2020-07-27 new2020/sections/sec01.Rmd

---
title: Section One
header-includes:
   - \usepackage{cleveref}
---

# Introduction

```{=latex}
\begin{equation}
\label{eq:no}
y = 2x
\end{equation}
```

Predicting bus arrival is ... quite something \( \cite{Elliott_2020} \)

I can just type things here, like ... yeah!
```{=latex}
\begin{equation}
\label{eq:hello}
x = \int 2x\; \mathrm{d}x
\end{equation}
```

We see in `\cref{eq:hello}`{=latex} that ... blah.