### 2020-07-27 new2020/sections/sec04.Rmd
Get subfiles working

---
title: Section Four
header-includes:
   - \usepackage{cleveref}
---

# Journey planning applications

* CDF makes it possible to answer many (often complex) jouney planning questions
* P(catch)
* P(arrive on time)
* P(transfer)
* this is a simple computation - can be done client side (i.e., on a user's phone) by passing CDF (small size, as e.g., JSON)