### 2020-07-28 new2020/sections/sec01.Rmd
Better build system

---
title: Section One
header-includes:
   - \usepackage{cleveref}
---

# Introduction

* the problem of predicting bus arrival - it needs to use real-time traffic information @cite:cn;
* however, many deployed methods are either specific to a provider/city, or don't make use of real-time data (only vehicle position and/or arrival delays, e.g., in Auckland)
* since the only logical source of ``traffic data'' in this setting is the transit vehicles themselves, makes sense to develop framework that uses them to estimate real-time network state @cite:Elliott_2020;
* in this example, particle filter is used to obtain a sample of points from the arrival time distribution - many many points, which cannot possibly be distributed or stored efficiently
* we propose a method of reducing this to a simple discrete CDF of arrival time (in minutes)