### 2020-09-29 new2020/sections/sec01.Rmd
begin adding structure

---
title: Section One
header-includes:
   - \usepackage{cleveref}
---

# Introduction

* the problem of predicting bus arrival - it needs to use real-time traffic information @cite:cn;
* however, many deployed methods are either specific to a provider/city, or don't make use of real-time data (only vehicle position and/or arrival delays, e.g., in Auckland)
* since the only logical source of "traffic data" in this setting is the transit vehicles themselves, makes sense to develop framework that uses them to estimate real-time network state @cite:Elliott_2020;
* in this example, particle filter is used to obtain a sample of points from the arrival time distribution - many many points, which cannot possibly be distributed or stored efficiently
* we propose a method of reducing this to a simple discrete CDF of arrival time (in minutes)
* we show that these results can be used to answer some common journey planning questions, in real-time, which outperform current "method"

Some of the references include:

* @cite:Cathey_2003;
* @cite:Yu_2006,Yu_2010,Yu_2011;
* @cite:Hans_2015;