### 2020-09-29 new2020/sections/sec02.Rmd

---
title: Section Two
header-includes:
   - \usepackage{cleveref}
---

# Background

Before describing the process of obtaining arrival time distributions, we must first define the framework with which we obtain vehicle and network state estimates. @cite:Elliott_2020; present a process for constructing a transit road network from raw GTFS data.

## A transit road network

* information about GTFS
* the concept of converting it into a network (as per @cite:Elliott_2020;)
* end product is real-time estimates of traffic state

## Estimating network and vehicle state using particle filtering

* particle filter on vehicle state
* used to estimate vehicle speed, position
* obtain distribution of travel times for each vehicle along each road