### 2020-12-02 new2020/sections/sec01.Rmd
intro

---
title: Section One
header-includes:
   - \usepackage{cleveref}
---

# Introduction

Public transport is an essential service for any large city, and key to making it successful is its reliability. While extreme and costly measures exist---such as building new infrastructure---a simpler, cost-effective approach is to simply provide reliable real-time information. This makes it easier for commuters to use and should hopefully improve ridership.

Predicting bus arrival times is nothing new---indeed, many unique methods have been presented and deployed around the world. The difficulty remains, however, in accounting for real-time traffic conditions in a reliable and generalised way.

@citet:Dailey_2001; used historical travel time data to predict a vehicle's arrival time based on its current location, while @citet:Jeong_2005; used a Kalman filter to model real-time travel times along links of a route. Other examples, such as @citet:Yu_2006,Yu_2010,Yu_2011; have used machine learning models and some *route pooling* to estimate traffic conditions along specific high-flow routes. Even more recently, taxi data was used by @citet:cn; to model real-time traffic, which was then used to predict arrival times. The concensus in the literature is that reliable predicts depend on real-time traffic estimates, however the methods used to do so have been varied and not generally applicable to services outside the target location.

In many cities, including out test locatio of Auckland, New Zealand, only have access to real-time transit vehicle location data, but no means of observing traffic conditions in real-time. @citet:Elliott_2020; proposed a simple method of constructing a *transit network* from GTFS @citep:GoogleDevelopers_2006; data, and using this along with real-time data to estimate real-time traffic state. In their paper, @citeauthor:Elliott_2020; used a particle filter to model a transit vehicle, as this has been shown @citep:Hans_2015; to be an effective method of estimating vehicle states given a wide range of plausible trajectories.

In this paper, we build on the work of @citet:Elliott_2020; and use the particle filter to estimate bus arrival times at stops using a real-time network state estimated from the buses themselves. Since the particle filter results in a large sample of estimates, we also propose a simplification of the CDF that can easily be distributed to users and used in journey planning applications.

<!--
* the problem of predicting bus arrival - it needs to use real-time traffic information @cite:cn;
* however, many deployed methods are either specific to a provider/city, or don't make use of real-time data (only vehicle position and/or arrival delays, e.g., in Auckland)
* since the only logical source of "traffic data" in this setting is the transit vehicles themselves, makes sense to develop framework that uses them to estimate real-time network state @cite:Elliott_2020;
* in this example, particle filter is used to obtain a sample of points from the arrival time distribution - many many points, which cannot possibly be distributed or stored efficiently
* we propose a method of reducing this to a simple discrete CDF of arrival time (in minutes)
* we show that these results can be used to answer some common journey planning questions, in real-time, which outperform current "method"

Some of the references include:

* @cite:Cathey_2003;
* @cite:Yu_2006,Yu_2010,Yu_2011;
* @cite:Hans_2015;
-->